---
title: 'Evaluating existing statistical methods for linear segmented relationship:
  A Simulation Study'
author: "Noah Degoulange"
date: "`r Sys.Date()`"
output:
  bookdown::html_book:
    theme: united
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE, fig.align='center'}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = "/indirect/student/noahdegoulange/github/mixedBreak")
```


```{r pkg, include=F}
library(dplyr)      # data manipulation
library(tidyr)
library(stringr)    # character manipulation
library(ggplot2)    # plots/graphs
library(readxl)     # read excel files
library(knitr)      # markdown layout
library(gt)         # table formatting
library(latex2exp)  # LaTeX in graphs
library(rsimsum)    # analyze/plot results from simulation study
```


```{r load_results, include=FALSE}
estimates.seglme <- read_excel("data/esimates_seglme.xlsx")
estimates.segreg <- read_excel("data/esimates_segreg.xlsx", guess_max = 10000L)
true.parameters <- read_excel("data/true_parameters.xlsx", sheet = "true.parameters")
scenario.data <- read_excel("data/true_parameters.xlsx", sheet = "scenario.data")
all.seeds <- read_excel("data/random_seed.xlsx", sheet = "all.seeds")$allseeds
# ending.seed <- read_excel("data/random_seed.xlsx", sheet = "ending.seed")$ending.seed
# restore with .GlobalEnv$.Random.seed <- ending.seed
TRUE.PSI <- 77
```


```{r compute_label, include=FALSE}
ref.ind <- which(scenario.data$fac.var=="REFERENCE")
scenario.data$fac.var[ref.ind] <- scenario.data$fac.var[ref.ind-1]
scenarLabel <- character(nrow(scenario.data))
for(i in 1:nrow(scenario.data)){
  scenarLabel[i] <- paste(scenario.data$fac.var[i],"=", 
                          scenario.data[i, scenario.data$fac.var[i]])
}
scenarLabel[ref.ind] <- paste(scenarLabel[ref.ind], "(REF)") # should I ?
reorder.index <- c(2,1,12,11,13,14,9,10, 3,5,6,7,8,4) # custom display order
scenario.data$scenarLabel <- as.factor(scenarLabel)
scenario.data$scenarLabel <- factor(scenario.data$scenarLabel, 
                                    levels(scenario.data$scenarLabel)[reorder.index])
levels(scenario.data$scenarLabel) <- paste0("(", 1:nrow(scenario.data), ") ", 
                                            levels(factor(scenario.data$scenarLabel)))
```



<!-- =========================================================== -->
<!-- =========================================================== -->


# Introduction



## Notations and definitions

In general, the hat $\hat$ notation will be reserved for estimates.

- A Data Generating Mechanism (DGM), or scenario
- 
- $\psi,\hat\psi_i$ resp. the true breakpoint value and its population-level estimate on dataset nb $i$,
- $\hat\psi_i$ the 
- $\hat\sigma_{\hat\psi}$ the empirical standard error (SE) of the breakpoint estimate
- The Monte Carlo Standard Error (Monte Carlo SE / MCSE) helps quantifying the uncertainty due to the finite number of repetitions done in the simulation study.




<!-- =========================================================== -->
<!-- =========================================================== -->



# Missing values

The first reported outcome are missing values counts and rates. We made a prior assumption of an expected 20% of missing estimates to derive the number of repetitions $n_{sim}$ for each scenario.

## _segmented.lme_

<!-- ### Overall rate -->

<!-- ```{r err_lme} -->
<!-- estimates.seglme %>%  -->
<!--   select(simID, error) %>% -->
<!--   distinct() %>%  -->
<!--   summarise(error.pct = round(100*sum(!is.na(error))/n(), 2), -->
<!--             error.nb = sum(!is.na(error))) %>% -->
<!--   gt() %>% -->
<!--   tab_header(title = "Missing estimates counts and rates", -->
<!--              subtitle = " Method 'segmented.lme'") %>% -->
<!--   opt_row_striping(row_striping = TRUE) -->
<!-- ``` -->

<!-- The overall rate of missing values for method _segmented.lme_ is 16.3%. On average, the number of missing values is below the one we expected (20%). However, scenario-wise, we have the following distribution: -->

### Scenario-wise rate

The distribution of missing estimates for every DGM is the following

```{r err_lme_scenar}
estimates.seglme %>% 
  select(simID, error) %>% 
  distinct() %>% 
  left_join(true.parameters[, c("simID", "scenarID")], by="simID") %>% 
  left_join(scenario.data[, c("scenarID", "scenarLabel")], by="scenarID") %>%
  distinct() %>%
  group_by(scenarLabel) %>%
  summarise(error.pct = round(100*sum(!is.na(error))/n(), 2),
            error.nb = sum(!is.na(error)),
            power.ok = (error.pct<20)) %>%
  gt(rowname_col = "scenarLabel") %>%
  tab_header(
    title = "Missing estimates counts and rates",
    subtitle = "Method 'segmented.lme' - Distribution per DGM (n.sim=594)"
  ) %>%
  tab_stubhead(label = "DGM") %>%
  opt_row_striping(row_striping = TRUE)
  # TODO - add red highlight if error rate too high ?
```

There is a very high disparities in terms of missing data depending on the running scenario. 4 out of the 14 scenarios do not have enough successful repetitions to conclude at the desired Monte Carlo power on the coverage probability.



### Error sources

Strictly speaking, two different reasons con be invoked for explaining _segmented.lme_'s missing values, albeit they are both related to a singularity during the model estimation:

```{r err_lme_reasons}
error.lme <- estimates.seglme %>% 
  select(simID, error) %>% 
  distinct() %>%
  select(error)
error.lme <- as.factor(error.lme$error)
# levels(error.lme) # aggregate identical error (up to a certain irrelevant value)
levels(error.lme) <- c("nlminb problem, singular convergence (7)", 
                       rep(substr(levels(error.lme)[2], 0, 34), nlevels(error.lme)-1))
knitr::kable(table(error.lme), align = 'll')
```

These errors came from the preliminary mixed model used for initializing the _segmented.lme_ method (CHECK IT). It indeed is a weakness of the method, namely because the initial mixed model is always miss-specified (no breakpoint is taken into account initially).


## _segreg_

<!-- ### Overall rate -->

<!-- ```{r err_segreg} -->
<!-- estimates.segreg %>%  -->
<!--   select(simID, error) %>%  -->
<!--   distinct() %>%  -->
<!--   summarise(error.pct = round(100*sum(!is.na(error))/n(), 2), -->
<!--             error.nb = sum(!is.na(error))) %>% -->
<!--   gt() %>% -->
<!--   tab_header( -->
<!--     title = "Overall missing values count and rate", -->
<!--     subtitle = "Method 'segreg'" -->
<!--   ) %>% -->
<!--   opt_row_striping(row_striping = TRUE) -->
<!-- ``` -->

<!-- The number of missing estimates is mostly reasonable for method _segreg_, with an overall rate of 2.63%.  -->


### Scenario-wise rate

```{r err_segreg_scenar}
estimates.segreg %>% 
  select(simID, error) %>% 
  distinct() %>% 
  left_join(true.parameters[, c("simID", "scenarID")], by="simID") %>% 
  left_join(scenario.data[, c("scenarID", "scenarLabel")], by="scenarID") %>%
  distinct() %>%
  group_by(scenarLabel) %>%
  summarise(error.pct = round(100*sum(!is.na(error))/n(), 2),
            error.nb = sum(!is.na(error)),
            power.ok = (error.pct<20)) %>%
  gt(rowname_col = "scenarLabel") %>%
  tab_header(
    title = "Missing values count and rate per scenario",
    subtitle = "Method 'segreg'"
  ) %>%
  tab_stubhead(label = "DGM") %>%
  opt_row_striping(row_striping = TRUE)
```

Most unsurprisingly, the number of missing values remains below for every scenario in this case, ranging from 0.17% to at most 8.25%. Since the model is patient-wise, the more patient in a dataset, the more errors could happen for this method, so the higher rate of missing values for scenario $n_{obs}=100$ is just an artifact rather than actual limitations at the individual level.

In particular, the convergence of the model is sensible to the starting values of the breakpoint With real data, in case it does not work in the first try, one could always visually inspect the scatter plot to provide a better initial guess to the model.



### Error sources

4 different kinds of errors occured for this method:

```{r err_segreg_reasons}
error.segreg <- estimates.segreg %>% 
  select(simID, error) %>% 
  distinct() %>%
  select(error)
error.segreg <- as.factor(error.segreg$error)
# levels(error.segreg) # truncated ID for which error occurred
error.segreg2 <- factor(sub(".*==> ", "", error.segreg))
levels(error.segreg2) <- c(
  "$ operator is invalid for atomic vectors",
  "Not enough information for the fit: too many psi/small sample/replicated data",
  "psi values too close each other. Please change (decreases number of) starting values",
  "starting psi out of the range"
)
knitr::kable(table(error.segreg2), align = 'll') 
```

At least two errors that could have been prevented with a robust code:

1. $ operator is invalid for atomic vectors
2. starting psi outside of the range (of *time* variable)

However, these were not encountered during the test run.

## Comparison of both methods

The following table presents the joint distribution of missing estimates:

```{r err_compare}
error.distrib <- estimates.seglme %>%
  select(simID, error) %>% 
  distinct() %>% 
  inner_join(estimates.segreg %>% select(simID, error) %>% distinct(), by = join_by(simID)) %>%
  mutate(
    error.common = case_when(
      !is.na(error.x) & !is.na(error.y) ~ "both",
      !is.na(error.x) ~ "segmented.lme",
      !is.na(error.y) ~ "segreg",
      .default = "None"),
    fail.segmented.lme = if_else(!is.na(error.x), "Missing - segmented", "CV - segmented"), 
    fail.segreg = if_else(!is.na(error.y), "Missing - segreg", "CV - segreg")
  )
knitr::kable(table(error.distrib %>% select(starts_with("fail."))), align = rep('c', 2))
```

19 datasets produced missing estimates for both methods (<0.25%), suggesting that errors in both methods happened independently. Datasets produced under failed repetitions have not been examined.

<!-- ## TODO - Explore datasets from which an error has occurred -->
<!-- ```{r err_explore} -->
<!-- error.seed <- all.seeds[error.distrib$simID] -->
<!-- ``` -->



<!-- =========================================================== -->
<!-- =========================================================== -->



# Graphical exploratory analysis

Before undertaking formal analysis of the performance measurements, graphical exploratory analysis will be helpful in assessing some characteristics of the simulated datasets, namely: the breakpoint time-distribution, correlation and systematic differences between methods, and issues with coverage.

## Distribution of $\hat{\psi_i}$ for each method and scenario

This graph aims at having an overview of the breakpoint estimates distribution, especially looking for outliers. The parameter of interest is $\psi$, the breakpoint time-location, and was simulated following a truncated log-normal distribution, with mean $\mu_\psi=77$ minutes for every data-generating mechanism (DGM, will be equally referred to as scenario).


### _segmented.lme_

The distribution of the estimated breakpoints $\hat\psi$ for the reference scenario is plotted below:

```{r est_dist.lme, warning=FALSE}
dataset1 <- estimates.seglme %>% 
  select(simID, break.avg) %>%
  distinct() %>% 
  left_join(true.parameters[, c("simID", "scenarID")], by="simID") %>% 
  left_join(scenario.data[, c("scenarID", "scenarLabel")], by="scenarID") %>%
  distinct()

ggplot(dataset1 %>% filter(str_detect(scenarLabel, "REF")), aes(x = break.avg, y = -0.5)) +
  geom_boxplot() +
  geom_density(aes(x = break.avg), inherit.aes = FALSE) +
  geom_vline(xintercept = TRUE.PSI, colour = "orange2", linetype = "dashed") +
  labs(title=TeX("Estimated \\hat{\\psi} distribution for reference scenario"),
       subtitle = TeX(paste("True mean \\psi = 77 ; Mean of estimates \\bar{\\hat{\\psi}} =",
                            round(mean(dataset1$break.avg[str_detect(scenarLabel, "REF")],
                                       na.rm=T),1), "; Median(\\hat{\\psi}) =",
                            round(median(dataset1$break.avg[str_detect(scenarLabel,"REF")],
                                                   na.rm=T),1)))) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

The distribution of estimated breakpoint is notably right-skewed, with the true mean being over the mode and median by 3.4 minutes, which could be a significant bias in the method. 
<!-- From a clinical perspective, is this amount of time relevant ?? -->

Most importantly, the scenario-wise distribution is the following:

```{r est_dist.lme.dgm, warning=FALSE}
dataset1.lme <- estimates.seglme %>% 
  select(simID, break.avg) %>% 
  distinct() %>% 
  left_join(true.parameters[, c("simID", "scenarID")], by="simID") %>% 
  left_join(scenario.data[, c("scenarID", "scenarLabel")], by="scenarID") %>%
  distinct()

# boxplot
ggplot(dataset1.lme , aes(x = scenarLabel, y = break.avg)) +
  geom_boxplot() +
  geom_hline(yintercept = TRUE.PSI, colour = "orange2", linetype = "dashed") +
  coord_flip() +
  labs(title=TeX("Empirical \\hat{\\psi} distribution for every data-generating mechanism"),
       subtitle = TeX("orange line is the target: simulated mean of breakpoint $\\psi=77$")) +
  ylab(TeX("Breakpoint estimates \\hat{\\psi} distribution")) +
  xlab("(scenario.nb) parameter varying from reference")
```

The labels on the _y_-axis corresponds to the parameter varying with all of the others being of the reference "cased-base" (scenario 7).

This panel reflects that the right skewness is dispatched among every DGM, though is it highly tempered by the longest plateau duration ($T_{peak}=143$ minutes).
The two notable outliers occurred for the same DGM, which has the highest residual variance ($\sigma_\varepsilon=2$) during slopes phases.

For the mixed model method, it visually looks like high standard deviation on the breakpoint $\sigma_\psi$ and low duration of plateau $T_{peak}$ biases the population-level estimate (under-estimation).

This observation is consistent with our hypothesis: because the plateau is not enforced into the mixed model, it might be that a positive slope is estimated in the second phase of the relationship between response and time (the true data being constant, it is mostly unlikely that the model estimates a negative slope). This additional degree of freedom induces bias by underestimating the breakpoint time coordinate $\hat{\psi}$.

TODO - REPORT SECOND PHASE SLOPES



### _segreg_

```{r est_dist.segreg.dgm, warning=FALSE}
dataset1.segreg <- estimates.segreg %>% 
  group_by(simID) %>%
  summarise(
    break.avg = mean(break.x1),
  ) %>%
  select(simID, break.avg) %>% 
  distinct() %>% 
  left_join(true.parameters[, c("simID", "scenarID")], by="simID") %>% 
  left_join(scenario.data[, c("scenarID", "scenarLabel")], by="scenarID") %>%
  distinct()

# boxplot
ggplot(dataset1.segreg , aes(x = scenarLabel, y = break.avg)) +
  geom_boxplot() +
  geom_hline(yintercept = TRUE.PSI, colour = "orange2", linetype = "dashed") +
  coord_flip() +
  labs(title=TeX("Empirical \\hat{\\psi} distribution for every data-generating mechanism"),
       subtitle = TeX("orange line is the target: simulated mean of breakpoint $\\psi=77$")) +
  ylab(TeX("Breakpoint estimates \\hat{\\psi} distribution")) +
  xlab("(scenario.nb) parameter varying from reference")
```

Overall, the distribution appears to be centered around the true value, albeit some scenario exhibits substantial deviation from the true value $\psi$. No remarkable outliers are to be noted.

## Bivariate plot of $\hat{\psi_i}$ for each DGM

With this plot, the goal is to try and look for correlation between the two methods, and any systematic differences for a given DGM.

The estimated breakpoint $\psi$ is represented on both axis, each axis corresponding to one of the two methods.

```{r biv_plot, warning=FALSE}
dataset2 <- estimates.segreg %>%
  group_by(simID) %>%
  summarise(est.psi.segreg = mean(break.x1)) %>%
  left_join(estimates.seglme %>%
              select(simID, break.avg) %>%
              distinct(), by = "simID") %>%
  rename(est.psi.seglme = break.avg) %>%
  left_join(true.parameters %>% select(simID, scenarID) %>% distinct(), by="simID") %>%
  left_join(scenario.data[, c("scenarID", "scenarLabel")], by="scenarID")

ggplot(dataset2 %>% filter(str_detect(scenarLabel, "REF")), aes(x=est.psi.segreg, y=est.psi.seglme)) +
  geom_point(alpha=.5) +
  geom_abline(slope=1, lty = "dashed", colour="orange2") +
  labs(title=TeX("Bivariate plot of \\hat{\\psi} for the reference scenario"),
       subtitle = paste("Dashed line is y = x ; Correlation is",
                        round(cor(dataset2$est.psi.segreg[str_detect(scenarLabel, "REF")],
                                  dataset2$est.psi.seglme[str_detect(scenarLabel, "REF")],
                                  use = "pairwise.complete.obs"), 2))) +
  xlab(TeX("Estimated \\hat{\\psi} for 'segreg'")) +
  ylab(TeX("Estimated \\hat{\\psi} for 'segmented.lme'"))

cor.test(dataset2$est.psi.segreg[str_detect(scenarLabel, "REF")],
         dataset2$est.psi.seglme[str_detect(scenarLabel, "REF")])$conf.int
```

The data pairs come from the same repetition (ie, they are estimated in the same simulated dataset) and are compared to the line of equality $y=x$. The scatter plot for the reference scenario depicts disagreement between the two methods, correlation being as low as $\hat\rho \hspace{1mm} [CI95\%] = 0.45 \hspace{1mm} [0.37 - 0.52]$.

The bivariate plot is produced in a panel for every data-generating mechanism below:

```{r biv_plot.dgm, fig.width=10, warning=FALSE}
dataset2.scenar <- dataset2 %>%
  group_by(scenarLabel) %>%
  mutate(rho = cor(est.psi.segreg, est.psi.seglme, use = "na.or.complete"),
         scenarLabel = as.factor(paste(scenarLabel, "; corr =", round(rho, 2))))

ggplot(dataset2.scenar, aes(x=est.psi.segreg, y=est.psi.seglme)) +
  geom_point(alpha=.5) + 
  facet_wrap(~scenarLabel) +
  geom_smooth(method = lm, formula = y~x) +
  geom_abline(slope=1, lty = "dashed", colour="orange2") +
  labs(title=TeX("Bivariate plot of \\hat{\\psi} for every scenario"),
       subtitle = paste("Dashed line is y=x ; Average correlation is", 
                        round(cor(dataset2$est.psi.segreg, dataset2$est.psi.seglme,
                                  use = "pairwise.complete.obs"), 2))) +
  xlab(TeX("Estimated \\hat{\\psi} for 'segreg'")) +
  ylab(TeX("Estimated \\hat{\\psi} for 'segmented.lme'"))
```

This panel exhibits heterogeneous between-method correlation among the different scenarios, the lowest being 0.33 for the scenario with the highest standard deviation on breakpoint $\sigma_\psi=33$, and the highest being 0.77 for the one with the longest plateau duration $T_{peak}=143$.

Overall, the correlation in the breakpoint estimate between the two methods is low to moderate, but not satisfactory.


## Zip plot

A zip plot is a plot of confidence intervals (CI) fractionally ranked by $|z_i|$, with $z_i := \frac{(\hat{\psi_i} - \psi)}{\sigma_{\psi}}$. It is a mean of understanding issue with coverage.
Under or over coverage typically arises in one of two situations:
(i) Bias in the breakpoint estimate $\hat\psi$ ; 
(ii) Misestimated empirical SE $\hat\sigma_{\hat\psi}$ of breakpoint estimate. Exempt of bias, underestimating this quantity leads to undercoverage, and overestimating it leads to overcoverage.


### _segmented.lme_

```{r zip_plot.seglme, fig.width=10, warning=FALSE}
seglme.df <- estimates.seglme %>%
  group_by(simID) %>%
  summarise(break.avg = first(break.avg),
            break.x1.sd = first(break.x1.sd),
            break.x1.random = first(break.x1.random),
            break.CI95.low = first(break.CI95.low),
            break.CI95.up = first(break.CI95.up))
segreg.df <- estimates.segreg %>%
  group_by(simID) %>%
  summarise(break.avg = mean(break.x1),
            break.x1.sd = sd(break.x1),
            break.CI95.low = mean(break.CI95.low),
            break.CI95.up = mean(break.CI95.up))

# dataset with z.score for zip plots (TODO - patch z score ?)
dataset <- seglme.df %>%
  bind_rows(segreg.df, .id = "method") %>%
  mutate(method = if_else(method==1, "segmented.lme", "segreg")) %>%
  left_join(true.parameters %>% select(simID, scenarID) %>% distinct(),
            by = "simID") %>%
  left_join(scenario.data %>% select(scenarID, scenarLabel), by= "scenarID") # %>%
  # group_by(scenarLabel) %>%
  # mutate(z.score = (break.avg - TRUE.PSI)/break.x1.sd,
  #        z.rank = rank(abs(z.score))/n(),
  #        contains = TRUE.PSI > break.CI95.low & TRUE.PSI < break.CI95.up)
# 
# p <- ggplot(dataset %>% filter(method=="segmented.lme"),
#        aes(y = z.rank, x = break.CI95.low, xend = break.CI95.up,
#            colour = contains)) +
#   geom_segment() +
#   facet_wrap(~scenarLabel) +
#   geom_vline(xintercept = TRUE.PSI, linetype = "dashed") +
#   geom_hline(yintercept = 0.95, linetype = "dashed") +
#   labs(colour = "") +
#   xlab("95% confidence intervals") +
#   ylab("Fractional centiles of |z-score|") +
#   scale_colour_hue(labels = c("Non-coverers", "Coverers", NA)) +
#   theme(legend.position = "bottom")

seglme.res <- simsum(
  dataset %>% filter(method=="segmented.lme"),
  estvarname = "break.avg",
  se = "break.x1.sd",
  true = TRUE.PSI, # true value of estimated parameter
  by = "scenarLabel", # factor of data-generating mechanism
  ci.limits = c("break.CI95.low", "break.CI95.up"),
  x = TRUE # necessary for zip plots
)
# 
# cover.conf.int <- seglme.res$summ %>%
#   filter(stat == "cover") %>% 
#   mutate(cover.95.low = est + qnorm(0.025)*mcse,
#          cover.95.up = est - qnorm(0.025)*mcse) %>% 
#   group_by(scenarLabel)

# p +
#   geom_hline(data = cover.conf.int, aes(yintercept = cover.95.low), 
#              linetype = "dashed", linewidth = 1.2, colour = "yellow") + 
#   geom_hline(data = cover.conf.int, aes(yintercept = cover.95.up), 
#              linetype = "dashed", linewidth = 1.2, colour = "yellow") +
#   ylim(c(0.5, 1))

autoplot(seglme.res, type="zip") + # zip plot !
  labs(title = "Zip plot for method 'segmented.lme' for every DGM")
# TODO - is there a scale problem on the panel in which the |z-score| fraction 
# does not reach 1 ? I would say yes, fractional z should be computed for every DGM!
# FIX |z|-scale issue

# plot.data <- autoplot(seglme.res, type="zip")@data
# pp <- plot.data %>% 
#   select(scenarLabel:simID, z, lower:upper, rank) %>%
#   ggplot()
# pp +
#   geom_segment(aes(y=z.frac.rank, x=lower, xend=upper, colour=)) +
#   facet_wrap(~scenarLabel)
# # zip plot mystery...

```

Blue CIs are coverers (_ie_ they include the true value $\psi=77$), red CIs are non-coverers (_ie_ they do not include the true value $\psi$). The yellow dashed lines represents the confidence interval on the coverage probability estimate.

The panel shows substantial undercoverage for every DGM but  the first one (lowest $\sigma_\psi=5$). Non-coverers can take up to more than 50% of all CIs for a given DGM, which questions the legitimacy of their name.

Since we already know about the substantial bias on the breakpoint estimate $\hat\psi$, it becomes clear that the breakpoint SE was underestimated in most cases.

<!-- Ironically enough, whilst some parameters were expected to have a harmful impact on performances (such as lower number of observations or higher outlier probability), they tend to patch undercoverage due to the substantial underestimation in $\hat\psi$. -->


### _segreg_


```{r zip_plot.segreg, fig.width=10, warning=FALSE}
segreg.res <- simsum(
  dataset %>% filter(method=="segreg"), 
  estvarname = "break.avg", 
  se = "break.x1.sd",
  true = TRUE.PSI, # true value of estimated parameter 
  by = "scenarLabel", # factor of data-generating mechanism
  ci.limits = c("break.CI95.low", "break.CI95.up"),
  x = TRUE # necessary for zip plots
)
autoplot(segreg.res, type="zip") + # zip plot !
  labs(title = "Zip plot for method 'segreg' for every DGM")
# TODO - same scale fix todo (maybe in another file)
# autoplot(summary(seglme.res), type = "lolly", stats = "bias") # TODO - add 
```

On the other side, _segreg_ method demonstrates substantial overcoverage, with a point estimate being not significantly different from 100% for all DGM.
This is most likely due to an over-estimation of the variance of breakpoint $\hat\sigma_{\hat\psi}$. Scenario (2) exhibits a shift in $\hat\psi$ (potential overestimation) when its variance $\sigma^2_\psi$ is higher.



<!-- =========================================================== -->
<!-- =========================================================== -->



# Performance measurements and Monte Carlo Standard Errors

```{r metrics_dataset}
## creation of unique dataset with all results for both methods
seglme.res <- estimates.seglme %>%
  group_by(simID) %>%
  summarise(break.avg = first(break.avg),
            break.x1.sd = first(break.x1.sd),
            break.x1.random = first(break.x1.random),
            break.CI95.low = first(break.CI95.low),
            break.CI95.up = first(break.CI95.up))
segreg.res <- estimates.segreg %>% 
  group_by(simID) %>%
  summarise(break.avg = mean(break.x1),
            break.x1.sd = sd(break.x1),
            break.CI95.low = mean(break.CI95.low),
            break.CI95.up = mean(break.CI95.up))
dataset <- seglme.res %>%
  bind_rows(segreg.res, .id = "method") %>%
  mutate(method = if_else(method==1, "segmented.lme", "segreg")) %>%
  left_join(true.parameters %>% 
              select(simID, scenarID) %>% 
              distinct(), 
            by = "simID") %>%
  left_join(scenario.data %>%
              select(scenarID, scenarLabel), 
            by= "scenarID")

metrics <- simsum(
  dataset, 
  estvarname = "break.avg", 
  se = "break.x1.sd",
  true = TRUE.PSI, # true value of estimated parameter 
  methodvar = "method", # variable containing the methods to compare
  ref = "segreg",
  by = "scenarLabel", # factor of data-generating mechanism
  ci.limits = c("break.CI95.low", "break.CI95.up"),
  x = TRUE, # necessary for zip plots
  control = list(mcse = TRUE)
)

# every metrics in summary
summ <- summary(metrics, stats = c("thetamean", "se2mean", "bias", "empse", 
                                   "cover", "becover"))
```

Formal analysis of the simulation study results goes through performance measurements. In our case we decided to examine bias on the breakpoint estimate, empirical SE of the estimate, and coverage probability of confidence intervals.

The results are presented as point estimate along with their **Monte Carlo Standard Error (MCSE)**. This helps quantifying the uncertainty due to the limited number of repetitions we ran for every scenario ($n_{sim}=594$). Consequently, Monte Carlo confidence intervals can be derived for every performance measure estimate.



## Bias lollipop plot

```{r lolli_bias, fig.width=10}
autoplot(summ, stats = "bias", type = "lolly") +
  labs(title = TeX("Bias in breakpoint estimate \\hat{\\psi}"),
       subtitle = "For both methods and every DGM") +
  xlab(TeX("Relative bias in breakpoint estimate \\hat{\\psi}"))
```

One can only notice the substantial and significant bias that the mixed model suffers from. The plateau model performs much better. Though the latter demonstrates a significant bias in a lot of scenarios, it is often not clinically relevant quantity (12/14 have an absolute bias under 1 minutes).
This observation has to be tempered by the fact that we reported the population-level breakpoint $\hat\psi_i$. The individual breakpoints $\hat\psi_{ij}$ might demonstrates other properties for the same performance measurements.


<!-- ```{r lolli_cover} -->
<!-- # APPENDIX? -->
<!-- autoplot(summ, stats = "cover", type = "lolly") + -->
<!--   labs(title = TeX("Coverage probability for \\hat{\\psi}"), -->
<!--        subtitle = "For both methods and every DGM") + -->
<!--   xlab("Coverage probability estimate (CI95%)") -->
<!-- ``` -->



<!-- ```{r lolli_becover} -->
<!-- # APPENDIX - easier to comment with the numbers -->
<!-- autoplot(summ, stats = "becover", type = "lolly")  + -->
<!--   labs(title = TeX("Bias-eliminated coverage for \\hat{\\psi}"), -->
<!--        subtitle = "For both methods and every DGM") + -->
<!--   xlab("Bias eliminated coverage probability estimate (CI95%)") -->
<!-- ``` -->

We can assess the negative impact of bias on the coverage probability, though it clearly remains a substantial undercoverage, with an important variability across scenarios.



## Performance measurements table

```{r performance_table}
# computing the table of performance measurements
typeIerr <- .05
formatted.res <- summ$summ  %>%
  # select(-c("lower", "upper")) %>%
  filter(stat %in% c("bias", "empse", "cover", "becover")) %>%
  pivot_wider(names_from = method, values_from = c(est:mcse, lower:upper)) %>%
  arrange(stat) %>%
  rename(lower_seglme = lower_segmented.lme,
         upper_seglme = upper_segmented.lme) %>%
  mutate(
    round.lme = -floor(log10(abs(mcse_segmented.lme))),
    segmented.lme = paste0(round(est_segmented.lme, round.lme), " (", 
                           round(mcse_segmented.lme, round.lme), ")"),
    round.seg = -floor(log10(abs(mcse_segreg))),
    segreg = paste0(round(est_segreg, round.seg), " (", 
                    round(mcse_segreg, round.seg), ")"),
    bias_segreg.target = (lower_segreg < 0) & (0 < upper_segreg),
    cover_segreg.target = (lower_segreg < 0.95) & (0.95 < upper_segreg),
    bias_seglme.target = (lower_seglme < 0) & (0 < upper_seglme),
    cover_seglme.target = (lower_seglme < 0.95) & (0.95 < upper_seglme)
  )


# printing the table of performance measurements
formatted.res %>%
  select(stat, scenarLabel, segmented.lme, segreg) %>%
  rename(seglme = segmented.lme) %>%
  pivot_wider(names_from = stat, values_from = c("seglme", "segreg")) %>%
  relocate(seglme_bias, seglme_empse, seglme_cover) %>% # to order the table
  gt(
    rowname_col = "scenarLabel",
    row_group_as_column = TRUE
  ) %>%
  tab_header(title = "Performance measurements per method and DGM (MCSE in parenthesis)") %>%
  # tab_caption("values (MCSE) - green cells indicates that the target value is reached for the performance measure ") %>%
  tab_stubhead(label = md("**DGM**")) %>%
  cols_align(align = "left", columns = scenarLabel) %>%
  tab_spanner(label=md("**Bias**"), columns=ends_with("_bias")) %>%
  tab_spanner(label=md("**EmpSE**"), columns=ends_with("_empse")) %>%
  tab_spanner(label=md("**Coverage**"), columns=ends_with("_cover")) %>%
  tab_spanner(label=md("**Bias-eliminated coverage**"), columns=ends_with("_becover")) %>%
  tab_options(table.font.size = 12) %>%
  opt_stylize(style = 1, color = "cyan") %>%
  # conditional formatting on bias/coverage significance w\ respect to target (0 / .95) ?
  # 1. bias!=0
  tab_style(
    style = cell_fill(color = "#90eea8"),
    locations = cells_body(
      columns = "segreg_bias",
      rows = formatted.res$bias_segreg.target[formatted.res$stat=="bias"] )
  ) %>%
  tab_style(
    style = cell_fill(color = "#90eea8"),
    locations = cells_body(
      columns = "seglme_bias",
      rows = formatted.res$bias_seglme.target[formatted.res$stat=="bias"] )
  ) %>%
  # 2. cover!=0.95
  tab_style(
    style = cell_fill(color = "#90eea8"),
    locations = cells_body(
      columns = "segreg_cover",
      rows = formatted.res$cover_segreg.target[formatted.res$stat=="cover"] )
  ) %>%
  tab_style(
    style = cell_fill(color = "#90eea8"),
    locations = cells_body(
      columns = "seglme_cover",
      rows = formatted.res$cover_seglme.target[formatted.res$stat=="cover"] )
  ) %>%
  # 2. becover!=0.95
  tab_style(
    style = cell_fill(color = "#90eea8"),
    locations = cells_body(
      columns = "segreg_becover",
      rows = formatted.res$cover_segreg.target[formatted.res$stat=="becover"] )
  ) %>%
  tab_style(
    style = cell_fill(color = "#90eea8"),
    locations = cells_body(
      columns = "seglme_becover",
      rows = formatted.res$cover_seglme.target[formatted.res$stat=="becover"] )
  )
```

<center>
*point estimate (MCSE) - green cells indicates that the target value (e.g. 0 for bias, 0.95 for coverage) is reached for the performance measure*
</center>

Overall, there is a significant **bias** in the breakpoint estimate $\hat\psi$, being notably underestimated by the mixed model, and misestimated in more than half DGM for the plateau model.


Empirical SE can be commented in the light of **coverage probability**, which is not reaching the nominal level of 95% for any of the methods, all DGM combined.
The plateau model shows an overcoverage, reaching a coverage of 100% in every DGM. Because it does not suffer from a substantial bias, it is most certainly due to an overestimation of the empirical SE $\hat\sigma_{\hat\psi}$.
On the other hand, the mixed model suffers from a substantial undercoverage. It may be due to both bias and underestimation of the breakpoints SE. 

To take bias into consideration, we can look at another performance measure, the "**bias-eliminated coverage**", which helps evaluating the estimators' performance under the assumption that they are unbiased:
The plateau model is still reaching the ceiling of 100% coverage in every scenario. It is most certainly a consequence of overestimating the breakpoint SE $\hat\sigma_\hat\psi$ because bias appears to be moderate.
The mixed model coverage probability reaches the nominal level for only 1 DGM out of 14. Though removing bias substantially improves the coverage probability, it is still not reaching a satisfactory level.

Below are listed the different **effects of simulation parameters** on the performance measurements:

- $\sigma_\psi$ (DGM 1,2 + ref): The higher the breakpoint individual variance, the more biased estimates are. This is also negatively impacting the coverage probability
- $T_{peak}$ (DGM 3,4 + ref): The higher the duration of plateau, the less biased the breakpoint estimate $\hat\psi$ is. The longest plateau is the one scenario for which both methods performs well, in terms of bias at least.
- $\sigma_\varepsilon$ (5, 6 + ref): The residual variance (noise on the point measurements in the datasets) impacts significantly and negatively bias. It does not induce a significant change in the bias-eliminated coverage.
- $p_{out}$ (DGM 7,8 + ref): The individual probability of having exactly one outlier in the time series makes both breakpoint estimate $\hat\psi$ and its SE $\hat\sigma_{\hat\psi}$ significantly higher
- $n_{obs}$ (DGM 9-14): Most surprisingly, when the number of observations (number of subjects per datasets) increases, it may induced a bias in the mixed model, though it may be an artifact of random sampling given the MCSE associated. For the plateau model however, it reduces the bias on the estimation. As expected, the empirical SE decreases when $n_{obs}$ increases. $n_{obs}$ increasing also impacts negatively the coverage probability, but seems to have no significant impact on the bias-eliminated coverage for the mixed model.
<!-- (bias increasing could be compensated by the breakpoint SE $\hat\sigma_\hat\psi$ decreasing). -->


# Discussion

The results of this simulation study might appear quite surprising in the light of previous knowledge brought by Muggeo. It is however mentioned that the algorithm described "always will converge in situations with low variance, provided that the break-point exists". Since the methods have not been stretched to the point where it breaks, this could be a plausible explanation of why they not yield accurate estimates.

TODO - To discuss: 

- Poor performance on the reference scenario makes the comparison a little awkward in the one at a time plan (since models performs poorly by default, one cannot judge the impact of a simulation parameters by the absolute value of the performance measurement but rather in comparison to the reference itself. A parameter might improve the models' performance whilst they could still be insufficient, only better than the reference) REPHRASE.
- Quantitative links with the clinically relevant parameters (slope up, peak duration)




